### 2024-02-23 sections/55-appendix-application-covid.tex
Add draft of first part on covid nowcasting

\section{Additional material on Section~\ref{sec:application-covid}}\label{sec:appendix-application-covid}

\begin{figure}
    \centering
    \begin{subfigure}[t]{.31\textwidth}
        \missingfigure{KDE for true and nowcasts of lag 1.}
        \caption{\ac{kde} for true values and nowcasts of lag 1.}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}[t]{.31\textwidth}
        \missingfigure{KDE for true and nowcasts of lag 7.}
        \caption{\ac{kde} for true values and nowcasts of lag 7.}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}[t]{.31\textwidth}
        \missingfigure{KDE for true and nowcasts of lag 14.}
        \caption{\ac{kde} for true values and nowcasts of lag 14.}
    \end{subfigure}
    \caption{\Ac{kde} for true values and nowcasts of lags 1, 7, and 14 to assess the distribution of values. \hl{AUSWERTUNG; 10 \% der Werte liegen innerhalb von ... um den Ursprung.}}
    \label{fig:app-covid-kde}
\end{figure}

\begin{table}
    \centering
    \begin{tabular}{l X}
        \toprule
        Abbreviation & Nowcasting hub key \\
        \midrule
        EPI & Epiforecasts-independent \\
        ILM & ILM-prop \\
        KIT & KIT-simple\_nowcast \\
        LMU & LMU\_StaBLab-GAM\_nowcast \\
        RIVM & RIVM-KEW \\
        RKI & RKI-weekly\_report \\
        SU & SU-hier\_bayes \\
        SZ & SZ-hosp\_nowcast\\
        ENS-MEAN & NowcastHub-MeanEnsemble\\
        ENS-MED & NowcastHub-MedianEnsemble\\
        \bottomrule
    \end{tabular}
    \caption{Matching of the abbreviation to the key in the nowcasting hub .Information on the models and references are listed in \textcite[][Table 1]{Wolffram2023}.}
    \label{tab:app-covid-models}
\end{table}