### 2024-03-21 sections/55-appendix-application-covid.tex

\section{Additional material on Section~\ref{sec:application-covid}}\label{sec:appendix-application-covid}

%\begin{figure}
%    \centering
%    \begin{subfigure}[t]{.48\textwidth}
%        \includegraphics{plots/covid_nowcast/20_kde_lag_1}
%        \caption{\ac{kde} for true values and nowcasts of lag 1.}
%    \end{subfigure}\hspace{0.01\textwidth}
%    \begin{subfigure}[t]{.48\textwidth}
%        \includegraphics{plots/covid_nowcast/20_kde_lag_7}
%        \caption{\ac{kde} for true values and nowcasts of lag 7.}
%    \end{subfigure}\hspace{0.01\textwidth}
%    \begin{subfigure}[t]{.48\textwidth}
%        \includegraphics{plots/covid_nowcast/20_kde_lag_14}
%        \caption{\ac{kde} for true values and nowcasts of lag 14.}
%    \end{subfigure}
%    \caption{\Ac{kde} for true values and nowcasts of lags 1, 7, and 14 to assess the distribution of values. \hl{AUSWERTUNG}. Exclusion areas based on the 10\%  quantile of absolute values are listed in Table~\ref{tab:app-covid-marginals}.}
%    \label{fig:app-covid-kde}
%\end{figure}

\begin{table}
    \centering
    \begin{tabular}{l l}
        \toprule
        Abbreviation & Nowcasting hub key \\
        \midrule
        EPI & Epiforecasts-independent \\
        ILM & ILM-prop \\
        KIT & KIT-simple\_nowcast \\
        LMU & LMU\_StaBLab-GAM\_nowcast \\
        RIVM & RIVM-KEW \\
        RKI & RKI-weekly\_report \\
        SU & SU-hier\_bayes \\
        SZ & SZ-hosp\_nowcast\\
        ENS-MEAN & NowcastHub-MeanEnsemble\\
        ENS-MED & NowcastHub-MedianEnsemble\\
        \bottomrule
    \end{tabular}
    \caption{Matching the abbreviation to the key in the nowcasting hub.
    Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
    \label{tab:app-covid-models}
\end{table}

\begin{table}
    \centering
    \input{plots/covid_nowcast/10_marginal_analysis}
    \caption{Analysis of the nowcast and true differences for the lags 1, 7, and 14 days.
    The column (1), $l=l$ shows the number of values greater than zero for lag $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
    \label{tab:app-covid-marginals}
\end{table}

\begin{table}
    \centering
    \begin{subtable}[t]{\textwidth}
        \input{plots/covid_nowcast/30_trending_ratios_lag_1}
    \caption{1 day.}
    \end{subtable}
    \begin{subtable}[t]{\textwidth}
        \input{plots/covid_nowcast/30_trending_ratios_lag_14}
        \caption{14 days.}
    \end{subtable}
    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the lag 1 and 14 days. The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values. }
    \label{tab:app-covid-trending-ratios-lag-1-14}
\end{table}

\begin{figure}
\centering
\includegraphics{plots/covid_nowcast/30_4q_plots}
\caption{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the lags of one, seven, and 14 days. The spread in both directions increases with the lag. }
\label{fig:app-covid-4q}
\end{figure}

\begin{figure}
    \centering
%    \includegraphics{}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
    \caption{Conditional trending plot for lag 1.}\label{fig:app-covid-cond-prob-1}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_14}
    \caption{Conditional trending plot for lag 1.}\label{fig:app-covid-cond-prob-14}
    \end{subfigure}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_1}
    \caption{Trending ratio over exclusion area size in $\diffx$ for lag 1.}\label{fig:app-covid-trending-ratio-1}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_14}
    \caption{Trending ratio over exclusion area size in $\diffx$ for lag 1.}\label{fig:app-covid-trending-ratio-14}
    \end{subfigure}
    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the lag seven days.}
    \label{fig:app-covid-cond-prob-trending-ratio-1-14}
\end{figure}