### 2024-07-28 CoverLetter/letter-to-the-editor.tex
Inital commit cover letter

\documentclass[%
 KITletter, %the letter class option
 %bwKITlogo, %--- activate to have the KIT logo black/white instead of coloured
 %officeprintableoutline, %--- activate for a layout printable on office printers
 %nooutlineframe, %--- activate tp print the letter on the pre-printed form of KIT
 %class options that do not change the layout can be added here, e.g. twoside,
fromemail,fromurl, %--- scrlttr2 options; activate as needed
 USenglish,%UKenglish%,french%,ngerman (already default) ... %--- activate one (or more) language options as needed
 %helvet, %heros, %--- use one of these to have a fallback font as default
 %cursor %--- activate to have the fallback typewriter font as default
  backaddress=off,
foldmarks=off,
 ]{scrlttr2}
\usepackage{hyperref}%for detailed information see: texdoc scrguide

%--- the back address (placed above the adressee)
%\setkomavar{backaddress}{KIT \textbar{} Dep. of Economics and Management \textbar{} Kaiserstr. 89 \textbar{} Karlsruhe, Germany}
\setkomavar{backaddress}{}

%--- the possible fields for the sender information (right of the addressee):
     %-- remove the fields not needed or set them to "{}",
     %-- use the optional arguments to adapt the respective descriptions or even to divert fields, e.g.
     %   \setkomavar{myref}[Position]{Project leader}
\setkomavar{fromname}{Jonas Rieger}
%\setkomavar{headofOU}{Prof. Vorname Surename}
\setkomavar{fromaddress}{Karlsruhe Institute of Technology\\Institute of Operations Research\\Chair Prof. Dr. O. Grothe – Analytics and Statistics\\76131 Karlsruhe, Germany}
%\setkomavar{fromphone}{0721-608-xxxxx}          %(to be activated by class option)
%\setkomavar{fromfax}{0721-608-xxxxx}            %          "
\setkomavar{fromemail}[Email:]{jonas.rieger@kit.edu}%          "
\setkomavar{fromurl}[Web:]{https://as.ior.kit.edu/}      %          "
%\setkomavar{contact}{Vorname Surename}
%\setkomavar{myref}{xxxx}
%\setkomavar{fromlogo}{\includegraphics[width=17mm]{info}}%(to be activated by class option)
%(the info graphic comes from the standard package "notes"
% and must in case of need b replaced by your individual logo)

%--- set-up for the main part and the closing of the letter:
%\setkomavar{title}{Testtitel} %--- rarely needed, placed before the subject line
\setkomavar{subject}{Submission to BMC Medical Research Methodology} %--- placed before the opening phrase
%\setkomavar{signature}{\includegraphics[width=2.5cm]{unterschrift}\\Jonas Rieger} %--- used by \closing{...}; overwrites the fromname variable
\setkomavar{signature}{[signature to be inserted]\\Jonas Rieger} %--- used by \closing{...}; overwrites the fromname variable
                                             %(you can provide a hand-written signature with \includegraphics)
\setplength{sigbeforevskip}{\baselineskip} %--- skip between closing phrase and signature

\setkomavar{firstfoot}{}%activate to suppress the footer of the letterhead page

\begin{document}

%\justifying %--- activate to have justified instead of ragged-right text

\def\pagename{} %--- activate to suppress "Page", "Seite" etc.  before the page number on pages 2 onwards

\begin{letter}{%
To the Editor of \\ BMC Medical Research Methodology
%\\LAND
}

\opening{Dear Dr. Lo Monaco,}

We hereby submit our manuscript

\begin{quote}
\textit{The assessment of the ability of measurements, nowcasts, and forecasts to track changes}
\end{quote}

by Bolin Liu, Bernd Saugel, Oliver Grothe, and myself for your kind consideration for publication in BMC Medical Research Methodology.

The manuscript introduces novel statistical methods and visualization techniques for assessing the ability of measurements, nowcasts, and forecasts to predict the correct direction of change.
These methods are exemplified using three distinct medical and healthcare datasets.
Hence, the manuscript aligns directly with the scope of BMC Medical Research Methodology.

We declare that there are no potential competing interests related to this research or manuscript.
All authors have reviewed and approved this manuscript for submission.
We confirm that the content of this manuscript has not been published, nor has it been submitted for publication in another journal.
This work is original and has not been considered by any other journal.
Please note that the material is also part of my doctoral thesis.

We respectfully suggest Prof. Dr. Alexander Hapfelmeier (\url{alexander.hapfelmeier@mri.tum.de}) from the Institute of AI and Informatics in Medicine at the Technical University of Munich, and Dr. Simon Vistisen (\url{vistisen@clin.au.dk}), Associate Professor and Business Engagement Partner in the Department of Clinical Medicine at Aarhus University and the Department of Anaesthesia \& Intensive Care at Aarhus University Hospital, as potential reviewers for our manuscript.

We are confident this paper represents a valuable and novel contribution to the evaluation of forecasting, nowcasting, and measurement methods in healthcare research.

\closing{\iflanguage{ngerman}{Mit freundlichen Grüßen}{Yours sincerely}}
%Jonas Rieger
%another idea for the closing:
%\setkomavar{signature}{Vorname Nachname\\Position}
%\setplength{sigbeforevskip}{0mm}
%\closing{\rule{12em}{0.5bp}}

\end{letter}

\end{document}